### 2022-11-10 R/model_fcs_target_markdown.Rmd
Update FCS model 2

---
title: "Favourable Conservation Status (FCS): <br> Analysis of Bauer et al. (unpublished) Field experiment"
author: "Markus Bauer <br>"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

#### Packages

```{r message = FALSE, tidy = TRUE}
library(here)
library(tidyverse)
library(ggbeeswarm)
library(patchwork)
library(brms)
library(DHARMa)
library(bayesplot)
library(loo)
library(tidybayes)
library(emmeans)

```

```{r echo = FALSE}
rm(list = ls())
rm(list = setdiff(ls(), c(
  "m_simple", "m_full", "m1", "m3", "m3_flat", "m_simple_flat", "m3_iter"
  )))
```

#### Load data

```{r}
sites <- read_csv(here("data", "processed", "data_processed_sites.csv"),
                  col_names = TRUE, na = c("na", "NA", ""), col_types =
                    cols(
                      .default = "?",
                      plot = "f",
                      site = "f",
                      sand_ratio = "f",
                      substrate_depth = "f",
                      target_type = col_factor(levels = c(
                        "dry_grassland", "hay_meadow"
                        )),
                      seed_density = "f",
                      exposition = col_factor(levels = c(
                        "north", "south"
                      )),
                      survey_year = "d"
                    )) %>%
  ### Exclude data of seed mixtures
  filter(survey_year != "seeded") %>%
  mutate(
    survey_year_fct = factor(survey_year),
    botanist_year = str_c(survey_year, botanist, sep = " "),
    botanist_year = factor(botanist_year),
    n = fcs_target,
    id = factor(id)
    ) %>%
  select(
    id, plot, site, exposition, sand_ratio, substrate_depth, target_type,
    seed_density, survey_year_fct, survey_year, botanist_year, n
    )
```

# Statistics

## Data exploration

### Graphs of raw data

```{r echo = FALSE}
plot1 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = sand_ratio)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Sand ratio [vol%] (Data of 2021)")
plot2 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = substrate_depth)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Substrate depth [cm] (Data of 2021)")
plot3 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = target_type)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Target type (Data of 2021)")
plot4 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = seed_density)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Seed density [g/m²] (Data of 2021)")
(plot1 + plot2) / (plot3 + plot4)
plot1 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = exposition)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Exposition (Data of 2021)")
plot2 <- ggplot(sites, aes(y = n, x = survey_year_fct)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Survey year")
plot3 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = site)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Blocks (Data of 2021)")
plot4 <- ggplot(sites, aes(y = n, x = botanist_year)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Botanists and survey year") +
  theme(axis.text.x = element_text(angle = 90))
(plot1 + plot2) / (plot3 + plot4)
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = substrate_depth)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Substrate depth [cm] vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = seed_density, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. Seed density [g/m²]")
```

### Outliers, zero-inflation, transformations?

```{r}
dotchart((sites$n), groups = factor(sites$exposition),
         main = "Cleveland dotplot")
sites %>% group_by(exposition) %>% count(site)
boxplot(sites$n)
plot(table((sites$n)), type = "h",
     xlab = "Observed values", ylab = "Frequency")
ggplot(sites, aes(x = n)) + geom_density()
ggplot(sites, aes(sqrt(n))) + geom_density()
```