### 2022-11-11 R/model_fcs_markdown.Rmd
Update model_fcs_markdown.Rmd

---
title: "Favourable Conservation Status (FCS): <br> Analysis of Bauer et al. (unpublished) Field experiment"
author: "Markus Bauer <br>"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

#### Packages

```{r message = FALSE, tidy = TRUE}
library(here)
library(tidyverse)
library(ggbeeswarm)
library(patchwork)
library(brms)
library(DHARMa)
library(bayesplot)
library(loo)
library(tidybayes)
library(emmeans)

```

```{r echo = FALSE}
rm(list = ls())
rm(list = setdiff(ls(), c(
  "m_simple", "m_full", "m1", "m3", "m3_flat", "m_simple_flat", "m3_iter"
  )))
```

#### Load data

```{r}
sites <- read_csv(here("data", "processed", "data_processed_sites.csv"),
                  col_names = TRUE, na = c("na", "NA", ""), col_types =
                    cols(
                      .default = "?",
                      plot = "f",
                      site = "f",
                      sand_ratio = "f",
                      substrate_depth = "f",
                      target_type = col_factor(levels = c(
                        "dry_grassland", "hay_meadow"
                        )),
                      seed_density = "f",
                      exposition = col_factor(levels = c(
                        "north", "south"
                      )),
                      survey_year = "d"
                    )) %>%
  ### Exclude data of seed mixtures
  filter(survey_year != "seeded") %>%
  mutate(
    survey_year_fct = factor(survey_year),
    botanist_year = str_c(survey_year, botanist, sep = " "),
    botanist_year = factor(botanist_year),
    n = fcs_target,
    id = factor(id)
    ) %>%
  select(
    id, plot, site, exposition, sand_ratio, substrate_depth, target_type,
    seed_density, survey_year_fct, survey_year, botanist_year, n
    )
```

# Statistics

## Data exploration

### Graphs of raw data

```{r echo = FALSE}
plot1 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = sand_ratio)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Sand ratio [vol%] (Data of 2021)")
plot2 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = substrate_depth)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Substrate depth [cm] (Data of 2021)")
plot3 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = target_type)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Target type (Data of 2021)")
plot4 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = seed_density)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Seed density [g/m²] (Data of 2021)")
(plot1 + plot2) / (plot3 + plot4)
plot1 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = exposition)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Exposition (Data of 2021)")
plot2 <- ggplot(sites, aes(y = n, x = survey_year_fct)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Survey year")
plot3 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = site)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Blocks (Data of 2021)")
plot4 <- ggplot(sites, aes(y = n, x = botanist_year)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Botanists and survey year") +
  theme(axis.text.x = element_text(angle = 90))
(plot1 + plot2) / (plot3 + plot4)
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = substrate_depth)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Substrate depth [cm] vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = seed_density, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. Seed density [g/m²]")
```

### Outliers, zero-inflation, transformations?

```{r}
sites %>% group_by(exposition) %>% count(site)
boxplot(sites$n)
ggplot(sites, aes(x = exposition, y = n)) + geom_quasirandom()
ggplot(sites, aes(x = n)) + geom_histogram(binwidth = 0.03)
ggplot(sites, aes(x = n)) + geom_density()
```

## Model building

### Models

Specifications for the models

```{r}
iter = 10000
chains = 4
thin = 2
priors <- c(
  set_prior("normal(0, 1)", class = "b"),
  set_prior("normal(0.1, 1)", class = "b", coef = "sand_ratio25"),
  set_prior("normal(0.2, 1)", class = "b", coef = "sand_ratio50"),
  set_prior("normal(0.1, 1)", class = "b", coef = "expositionsouth"),
  set_prior("normal(0.1, 1)", class = "b", coef = "survey_year_fct2019"),
  set_prior("normal(0.2, 1)", class = "b", coef = "survey_year_fct2020"),
  set_prior("normal(0.3, 1)", class = "b",coef = "survey_year_fct2021"),
  set_prior("cauchy(0, 1)", class = "sigma")
)
```

Model caluclations

```{r}
load(file = here("data", "processed", "model_fcs_1.Rdata"))
load(file = here("data", "processed", "model_fcs_2.Rdata"))
load(file = here("data", "processed", "model_fcs_3.Rdata"))
```

### Model comparison

```{r}
m_1 <- m1
m_2 <- m3
m_1$formula
m_2$formula
```

Conditional and marginal R² values

```{r}
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot) + (1 | botanist_year)) 
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot) + (1 | botanist_year)) 
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
```

### Model check

#### DHARMa

```{r}
createDHARMa(
  simulatedResponse = t(posterior_predict(m_1)),
  observedResponse = sites$n,
  fittedPredictedResponse = apply(t(posterior_epred(m_1)), 1, mean),
  integerResponse = TRUE
  ) %>%
  plot()
createDHARMa(
  simulatedResponse = t(posterior_predict(m_2)),
  observedResponse = sites$n,
  fittedPredictedResponse = apply(t(posterior_epred(m_2)), 1, mean),
  integerResponse = TRUE
  ) %>%
  plot()
```

#### Preparation

```{r}
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth30",
      "b_target_typehay_meadow",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sd_botanist_year__Intercept",
      "sigma"
    )
  )
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth30",
      "b_target_typehay_meadow",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sd_botanist_year__Intercept",
      "sigma"
    )
  )
hmc_diagnostics1 <- nuts_params(m_1)
hmc_diagnostics2 <- nuts_params(m_2)
y <- sites$n
yrep1 <- posterior_predict(m_1, draws = 500)
yrep2 <- posterior_predict(m_2, draws = 500)
loo1 <- loo(m_1, save_psis = TRUE, moment_match = FALSE)
loo2 <- loo(m_2, save_psis = TRUE, moment_match = FALSE)
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

#### Samling efficency/effectiveness (Rhat and EFF)

```{r}
range(draws1$rhat)
range(draws2$rhat)
range(draws1$ess_bulk)
range(draws2$ess_bulk)
range(draws1$ess_tail)
range(draws2$ess_tail)
```

#### MCMC diagnostics

```{r}
mcmc_trace(posterior1, np = hmc_diagnostics1)
mcmc_trace(posterior2, np = hmc_diagnostics2)
```

#### Posterior predictive check

##### Kernel density

```{r}
p1 <- ppc_dens_overlay(y, yrep1[1:50, ])
p2 <- ppc_dens_overlay(y, yrep2[1:50, ])
p1 / p2
ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$site)
ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$site)
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$exposition)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$exposition)
p1 / p2
ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$survey_year_fct)
ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$survey_year_fct)
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$target_type)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$target_type)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$seed_density)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$seed_density)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$sand_ratio)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$sand_ratio)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$substrate_depth)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$substrate_depth)
p1 / p2
```

##### Histograms of statistics skew

```{r}
p1 <- ppc_stat(y, yrep1, binwidth = 0.001)
p2 <- ppc_stat(y, yrep2, binwidth = 0.001)
p1 / p2
ppc_stat_grouped(y, yrep1, group = sites$site, binwidth = 0.001)
ppc_stat_grouped(y, yrep2, group = sites$site, binwidth = 0.001)
p1 <- ppc_stat_grouped(y, yrep1, group = sites$exposition, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$exposition, binwidth = 0.001)
p1 / p2
ppc_stat_grouped(y, yrep1, group = sites$survey_year_fct, binwidth = 0.001)
ppc_stat_grouped(y, yrep2, group = sites$survey_year_fct, binwidth = 0.001)
p1 <- ppc_stat_grouped(y, yrep1, group = sites$target_type, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$target_type, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$seed_density, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$seed_density, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$sand_ratio, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$sand_ratio, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$substrate_depth, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$substrate_depth, binwidth = 0.001)
p1 / p2
```

##### LOO-PIT plots

```{r}
p1 <- ppc_loo_pit_overlay(y, yrep1, lw = weights(loo1$psis_object))
p2 <- ppc_loo_pit_overlay(y, yrep2, lw = weights(loo2$psis_object))
p1 / p2
plot(loo1)
plot(loo2)
```

#### Autocorrelation check

```{r}
mcmc_acf(posterior1, lags = 10)
mcmc_acf(posterior2, lags = 10)
```

## Output of choosen model

### Model output

```{r}
prior_summary(m_1)
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot) + (1 | botanist_year)) 
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot) + (1 | botanist_year)) 
bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
draws1
mcmc_intervals(
  posterior1,
  prob = 0.5,
  prob_outer = 0.95,
  point_est = "mean"
)
sjPlot::plot_model(m_1, m_2, type = "est")
sjPlot::plot_model(m_2, type = "est")
sjPlot::plot_model(
  m_1, m_2, type = "est", ppd = TRUE,
  terms = c(
    "sand_ratio", "substrate_depth", "target_type",
    "seed_density",  "exposition", "survey_year_fct"
    )
  )
```

### Effect sizes

```{r}
(emm <- emmeans(m_1, revpairwise ~ target_type + sand_ratio |
                  exposition | survey_year_fct, type = "response"))
(emm <- emmeans(m_1, revpairwise ~ substrate_depth + sand_ratio |
                  exposition | survey_year_fct, type = "response"))
(emm <- emmeans(m_1, revpairwise ~ seed_density |
                  exposition | survey_year_fct, type = "response"))
```