### 2022-12-07 R/model_persistence_check.Rmd
Update model scripts

---
title: "Analysis of Bauer et al. (unpublished) Field experiment: <br> Persistence"
author: "<b>Markus Bauer</b> <br>"
date: "<b>`r format(Sys.time(), '%Y-%m-%d')`</b>"
output:
  github_document:
    toc: true
    toc_depth: 3
    dev: png
    fig_width: 7
    fig_height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  )
```

<b>Markus Bauer</b>

Technichal University of Munich, TUM School of Life Sciences, Chair of Restoration Ecology, Emil-Ramann-Straße 6, 85354 Freising, Germany

[markus1.bauer@tum.de](mailto:markus1.bauer@tum.de)

ORCiD ID: [0000-0001-5372-4174](https://orcid.org/0000-0001-5372-4174)
<br>
[Google Scholar](https://scholar.google.de/citations?user=oHhmOkkAAAAJ&hl=de&oi=ao)
<br>
GitHub: [markus1bauer](https://github.com/markus1bauer)

To compare different models, you only have to change the models in section 'Load models'

# Preparation

Favourable Conservation Status (FSC) sensu Helm et al. (2015) Divers Distrib [DOI: 10.1111/ddi.12285](https://doi.org/10.1111/ddi.12285)

Analysis motivated by Applestein et al. (2021) Restor Ecol [DOI: 10.1111/rec.13596](https://doi.org/10.1111/rec.13596)

Analysis guided by
<br>
<b>BARG</b> (Bayesian Analysis Reporting Guidelines): Kruschke (2021) Nat Hum Behav [DOI: 10.1038/s41562-021-01177-7](https://doi.org/10.1038/s41562-021-01177-7)
<br>
Model check: Gabry et al. (2019) J R Stat Soc A Stat [DOI: 10.1111/rssa.12378](https://doi.org/10.1111/rssa.12378)
<br>
Priors: Lemoine (2019) Oikos [DOI: 10.1111/oik.05985](https://doi.org/10.1111/oik.05985)
<br>
Model check: Deapoli & Schoot (2017) Psychol Methods [DOI: 10.1037/met0000065](https://doi.org/10.1037/met0000065)

#### Packages (BARG 2.A)

```{r libraries, message = FALSE}
library(here)
library(tidyverse)
library(ggbeeswarm)
library(patchwork)
library(brms)
library(DHARMa.helpers)
library(bayesplot)
library(loo)
library(tidybayes)
library(emmeans)
```

```{r echo = FALSE}
rm(list = ls())
```

#### Load data

```{r load-data}
sites <- read_csv(
  here("data", "processed", "data_processed_sites_temporal.csv"),
  col_names = TRUE, na = c("na", "NA", ""),
  col_types = cols(
    .default = "?",
    plot = "f",
    site = "f",
    sand_ratio = "f",
    substrate_depth = col_factor(levels = c("30", "15")),
    target_type = col_factor(levels = c("hay_meadow", "dry_grassland")),
    seed_density = "f",
    exposition = col_factor(levels = c("north", "south")),
    survey_year = "c"
  )
  ) %>%
  ### Exclude data of seed mixtures
  filter(presabu == "presence") %>%
  mutate(
    survey_year_fct = factor(survey_year),
    botanist_year = str_c(survey_year, botanist, exposition, sep = " "),
    botanist_year = factor(botanist_year),
    id = factor(id),
    n = persistence / 100
  ) %>%
  select(
    id, plot, site, exposition, sand_ratio, substrate_depth, target_type,
    seed_density, survey_year_fct, survey_year, botanist_year, n
  )
```

# Statistics

## Data exploration

### Graphs of raw data

```{r data-exploration, echo = FALSE}
plot1 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = sand_ratio)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Sand ratio [vol%]")
plot2 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = substrate_depth)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Substrate depth [cm]")
plot3 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = target_type)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Target type")
plot4 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = seed_density)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Seed density [g/m²]")
(plot1 + plot2) / (plot3 + plot4)
plot1 <- ggplot(sites %>% filter(survey_year == 2021),
                aes(y = n, x = exposition)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Exposition")
plot2 <- ggplot(sites, aes(y = n, x = survey_year_fct)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Survey year")
plot3 <- ggplot(sites %>% filter(survey_year == 2021), aes(y = n, x = site)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  facet_grid(~ survey_year_fct) +
  labs(title = "Blocks")
plot4 <- ggplot(sites, aes(y = n, x = botanist_year)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  labs(title = "Botanists and survey year") +
  theme(axis.text.x = element_text(angle = 90))
(plot1 + plot2) / (plot3 + plot4)
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = sand_ratio, y = n, color = substrate_depth)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Substrate depth [cm] vs. sand ratio [vol%]")
ggplot(data = sites,
       aes(x = seed_density, y = n, color = target_type)) + 
  geom_quasirandom(alpha = 0.5, dodge.width = 0.8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(exposition ~ survey_year_fct) +
  labs(title = "Target type vs. Seed density [g/m²]")
```

### Outliers, zero-inflation, transformations?

```{r outliers, echo = FALSE}
sites %>% group_by(exposition) %>% count(site)
boxplot(sites$n)
ggplot(sites, aes(x = exposition, y = n)) + geom_quasirandom()
ggplot(sites, aes(x = n)) + geom_histogram(binwidth = 0.03)
ggplot(sites, aes(x = n)) + geom_density()
```

## Models

### Load models (BARG 6.G)

Only here you have to modify the script to compare other models

```{r load-models, collapse = TRUE, message = FALSE}
load(file = here("outputs", "models", "model_persistence_2.Rdata"))
load(file = here("outputs", "models", "model_persistence_full.Rdata"))
load(file = here("outputs", "models", "model_persistence_2_prior.Rdata"))
# BARG 5.A/B/C
#load(file = here("outputs", "models", "model_persistence_2_flat.Rdata"))
m_1 <- m2
m_2 <- m_full
m_prior <- m2_prior
```

### Model specifications (BARG 1.D)

```{r formulas, collapse = TRUE}
m_1$formula
m_2$formula
```

```{r families, collapse = TRUE}
m_1$family
m_2$family
```

Amount of chains for MCMC

```{r model-specification-chains, collapse = TRUE}
m_1$fit@sim$chains
m_2$fit@sim$chains
```

Total amount of iterations for MCMC

```{r model-specification-iterations, collapse = TRUE}
m_1$fit@sim$iter
m_2$fit@sim$iter
```

Amount of iterations before burn-in

```{r model-specification-warmup, collapse = TRUE}
m_1$fit@sim$warmup
m_2$fit@sim$warmup
```

Thinning rate

```{r model-specification-thinning, collapse = TRUE}
m_1$fit@sim$thin
m_2$fit@sim$thin
```

### Preparation for analysis

```{r analysis-preparation}
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
  )
posterior_prior <- m_prior %>%
  posterior::as_draws() %>%
  posterior::subset_draws(
    variable = c(
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021"
    )
  )
hmc_diagnostics1 <- brms::nuts_params(m_1)
hmc_diagnostics2 <- brms::nuts_params(m_2)
y <- sites$n
yrep1 <- brms::posterior_predict(m_1, draws = 500)
yrep2 <- brms::posterior_predict(m_2, draws = 500)
loo1 <- brms::loo(m_1, save_psis = TRUE, moment_match = FALSE)
loo2 <- brms::loo(m_2, save_psis = TRUE, moment_match = FALSE)
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

### Priors (BARG 1.D/E)

#### Possible prior distributions

```{r possible-priors}
get_prior(n ~ target_type + exposition + sand_ratio + survey_year_fct +
            seed_density + substrate_depth +
            (1 | site/plot) + (1 | botanist_year),
          data = sites)
ggplot(data = data.frame(x = c(0, 1)), aes(x = x)) +
  stat_function(fun = dbeta, n = 101, args = list(shape1 = 3.5, shape2 = 2.2)) +
  expand_limits(y = 0) +
  ggtitle("Beta distribution for Intercept")
ggplot(data = data.frame(x = c(-.4, .4)), aes(x = x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = .2)) +
  expand_limits(y = 0) +
  ggtitle("Normal distribution for treatments")
ggplot(data = data.frame(x = c(-.4, .4)), aes(x = x)) +
  stat_function(fun = dcauchy, n = 101, args = list(location = 0, scale = 1)) +
  expand_limits(y = 0) +
  ggtitle("Cauchy distribution") # See Lemoine 2019 https://doi.org/10.1111/oik.05985
ggplot(data.frame(x = c(-.4, .4)), aes(x = x)) +
  stat_function(fun = dstudent_t, args = list(df = 3, mu = 0, sigma = 2.5)) +
  expand_limits(y = 0) +
  ggtitle(expression(Student~italic(t)*"-distribution")) # Software standard
```

#### Prior summary (BARG 1.D)

```{r prior-summary}
brms::prior_summary(m_1, all = FALSE)
```

#### Prior predictive check (BARG 1.E)

```{r prior-predictive-check}
bayesplot::mcmc_intervals(
  posterior_prior,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
  ) +
  theme_classic()
```

## Model check

### DHARMa

```{r dharma, collapse = TRUE}
DHARMa.helpers::dh_check_brms(m_1, integer = TRUE)
DHARMa.helpers::dh_check_brms(m_2, integer = TRUE)
```

### Sampling efficency and effectiveness (BARG 2.B/C)

#### Rhat (BARG 2.B)

```{r rhat, collapse = TRUE}
mcmc_rhat(draws1$rhat)
mcmc_rhat(draws2$rhat)
```

#### Effective sampling size (ESS) (BARG 2.C)

```{r ess, collapse = TRUE}
mcmc_neff(neff_ratio(m_1))
mcmc_neff(neff_ratio(m_2))
```

### MCMC diagnostics

#### Trace plots

```{r mcmc-trace, fig.height = 14}
mcmc_trace(posterior1, np = hmc_diagnostics1, facet_args = list(ncol = 2))
mcmc_trace(posterior2, np = hmc_diagnostics2, facet_args = list(ncol = 2))
```

#### Pairs plot

```{r mcmc-pairs, collapse = TRUE, fig.height = 6.5}
mcmc_pairs(m_1, off_diag_args = list(size = 1.2),
           pars = c(
             "b_sand_ratio25", "b_sand_ratio50", "b_substrate_depth15",
             "b_target_typedry_grassland", "b_seed_density8",
             "b_expositionsouth", "sigma"
           ))
mcmc_pairs(m_2, off_diag_args = list(size = 1.2),
           pars = c(
             "b_sand_ratio25", "b_sand_ratio50", "b_substrate_depth15",
             "b_target_typedry_grassland", "b_seed_density8",
             "b_expositionsouth", "sigma"
           ))
```

#### Parallel coordinate plot

```{r mcmc-parcoord, collapse = TRUE}
mcmc_parcoord(posterior1, np = hmc_diagnostics1) +
  theme(axis.text.x = element_text(angle = 45))
mcmc_parcoord(posterior2, np = hmc_diagnostics2) +
  theme(axis.text.x = element_text(angle = 45))
```

### Posterior predictive check

#### Kernel density

```{r kernel-density}
p1 <- ppc_dens_overlay(y, yrep1[1:50, ])
p2 <- ppc_dens_overlay(y, yrep2[1:50, ])
p1 / p2
ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$site)
ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$site)
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$exposition)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$exposition)
p1 / p2
ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$survey_year_fct)
ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$survey_year_fct)
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$target_type)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$target_type)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$seed_density)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$seed_density)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$sand_ratio)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$sand_ratio)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$substrate_depth)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$substrate_depth)
p1 / p2
```

#### Histograms of statistics skew

```{r histograms}
#p1 <- ppc_stat(y, yrep1, binwidth = 0.001)
#p2 <- ppc_stat(y, yrep2, binwidth = 0.001)
#p1 / p2
#ppc_stat_grouped(y, yrep1, group = sites$site, binwidth = 0.001)
#ppc_stat_grouped(y, yrep2, group = sites$site, binwidth = 0.001)
#p1 <- ppc_stat_grouped(y, yrep1, group = sites$exposition, binwidth = 0.001)
#p2 <- ppc_stat_grouped(y, yrep2, group = sites$exposition, binwidth = 0.001)
#p1 / p2
#ppc_stat_grouped(y, yrep1, group = sites$survey_year_fct, binwidth = 0.001)
#ppc_stat_grouped(y, yrep2, group = sites$survey_year_fct, binwidth = 0.001)
#p1 <- ppc_stat_grouped(y, yrep1, group = sites$target_type, binwidth = 0.001)
#p2 <- ppc_stat_grouped(y, yrep2, group = sites$target_type, binwidth = 0.001)
#p1 / p2
#p1 <- ppc_stat_grouped(y, yrep1, group = sites$seed_density, binwidth = 0.001)
#p2 <- ppc_stat_grouped(y, yrep2, group = sites$seed_density, binwidth = 0.001)
#p1 / p2
#p1 <- ppc_stat_grouped(y, yrep1, group = sites$sand_ratio, binwidth = 0.001)
#p2 <- ppc_stat_grouped(y, yrep2, group = sites$sand_ratio, binwidth = 0.001)
#p1 / p2
#p1 <- ppc_stat_grouped(y, yrep1, group = sites$substrate_depth, binwidth = 0.001)
#p2 <- ppc_stat_grouped(y, yrep2, group = sites$substrate_depth, binwidth = 0.001)
#p1 / p2
```

#### LOO (Leave one out)

```{r loo}
loo1
loo2
#plot(loo1)
#plot(loo2)
```

Leave one out probability integral transform

```{r loo-pit}
#p1 <- bayesplot::ppc_loo_pit_overlay(y, yrep1, lw = weights(loo1$psis_object))
#p2 <- bayesplot::ppc_loo_pit_overlay(y, yrep2, lw = weights(loo2$psis_object))
#p1 / p2

```

### Autocorrelation check

```{r autocorrelation, fig.height = 7}
mcmc_acf(posterior1, lags = 10)
mcmc_acf(posterior2, lags = 10)
```

## Model comparison

### Conditional <i>R</i>² values

```{r conditional-r2, collapse = TRUE}
brms::bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot)) 
brms::bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site/plot))
```

### Marginal <i>R</i>² values

```{r marginal-r2, collapse = TRUE}
brms::bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
brms::bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
```

### Bayes factor (BARG 3.C)

```{r bayes-factor}
brms::bayes_factor(m_1, m_2)
```

## Posterior distributions (BARG 3.B)

### Forest plot

Chosen model:

```{r posteriors-chosen-model}
mcmc_intervals(
  posterior1,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
  ) +
  theme_classic()
```

Second model:

```{r posterors-2nd-model}
bayesplot::mcmc_intervals(
  posterior2,
  prob = 0.66,
  prob_outer = 0.95,
  point_est = "mean"
  ) +
  theme_classic()
```

### Effect sizes

Effect sizes of chosen model just to get exact values of means etc. if necessary.

```{r effect-sizes}
#emmeans(m_1, revpairwise ~ target_type + sand_ratio | exposition |
#survey_year_fct, type = "response"))
```

```{r write-csv, echo = FALSE}
write.csv(draws1, here("outputs", "statistics", "table_persistence.csv"))
```

# Session info (BARG 2.A/6.A/6.B)

```{r session-info, echo = FALSE}
sessionInfo()
```