### 2023-03-02 manuscript.Rmd
Update manuscript.Rmd

---
title: "Fit by design: Developing substrate-specific seed mixtures for functional dike grasslands"
author: Markus Bauer*, Jakob Huber, Johannes Kollmann
output:
  officedown::rdocx_document:
    reference_docx: template_word.docx
csl: template_references_j_appl_ecol.csl
bibliography: bibliography.bib
---

```{r set-up, include = FALSE, message = FALSE}
library(here)
library(knitr)
library(officedown)
library(officer)
library(tidyverse)
library(flextable)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  dpi = 300,
  fig.cap = FALSE
  )
set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 11,
  text.align = "center"
  )
ft <- officer::fp_text(shading.color = "yellow")

word_spec <- function(x, prop = ft) ftext(text = toString(x), prop =  ft)
```

Restoration Ecology, TUM School of Life Sciences, Technical University of Munich, Germany

\* Corresponding author: [markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

\clearpage

\newpage

# Abstract {.unnumbered}

1.  Sowing is a well-established restoration technique to overcome dispersal limitation and site-specific seed mixtures are most effective to achieve functional communities. This is especially important if the restored vegetation has to protect critical infrastructure like roadsides and dikes. Here, an improved seed--substrate combination will secure slope stability, reduce mowing efforts, and generate biodiverse grasslands.
2.  A factorial field experiment addressed this topic on a dike at the Danube River in SE Germany in the 2018--2021 time period. On 288 plots, we tested three sand admixtures, two substrate depths, two seed densities and two seed mixture types (mesic hay meadow, semi-dry calcareous grassland) in north or south exposition, and measured the persistence of sown species, the Favourable Conservation Status (FCS), and the successional distance to reference sites as calculated from an ordination.
3.  Sand admixture had a slightly positive effect on target variables, while substrate depth, seed density and mixture type had no effects on species persistence or FCS.
4.  Overall, the sown vegetation developed in the desired direction, but a recovery debt remained after four years, and some plots still showed similarities to negative references from ruderal sites. In north exposition, hay meadow-seed mixtures developed closer to the respective reference communities than dry grasslands.
5.  In south exposition, the sown communities developed poorly which might be due to a severe drought during establishment and the addition and removal of a geotextile. This initial negative effect remained over the entire observation period.
6.  Synthesis and applications: Site-adapted seed mixtures make restoration more efficient. However, applying several seed--substrate combinations might foster beta diversity. Furthermore, additional management efforts are recommended, as they might be necessary to reduce the recovery debt and re-sowing after unfavourable conditions. \clearpage \newpage

# Keywords {.unnumbered}

Artificial soil mixture

Community assembly

Dry grasslands

Ecological restoration

Favourable Conservation Status

Hay meadow

Levee

Persistence

River embankment

Soil restoration

Sowing

Species composition

Species-rich grassland

\clearpage

\newpage

# Introduction

Grasslands can support an exceedingly high biodiversity and they provide several ecosystem services [@Dengler.2014; @Bardgett.2021]. However, they are globally endangered [@Bardgett.2021], and in Europe, calcareous grasslands and hay meadows are red-listed habitats [Category 3, 'vulnerable', @Janssen.2016]. Restoration is seen as a key factor to sustain biodiversity and ecosystem services [@COP.2014; @UnitedNations.2019b], and sowing is a well-established approach to establish species-rich grasslands [@Kiehl.2010]. Sowing high-diversity mixtures of local provenance produced by specialized companies is a promising way to scale up restoration efforts [@Freitag.2021], and to overcome the dispersal filters [@myers2009; @orrock2023]. However, there are still open questions about adjusting seed mixtures to specific site conditions and future climate conditions [@Torok.2021].

Restoration ecology can increase the predictability of restoration approaches [@Suding.2011; @Mouquet.2015] by using rigorous, repeatable, and transparent experiments based on advanced theory, which will finally strengthen evidence-based restoration [@Cooke.2018; @Wainwright.2018]. Local site conditions and the restoration method are key predictors for vegetation development after sowing [@Brudvig.2017b], and habitat and biotic filtering are the main assembly factors which can be manipulated by the choice of seed--substrate combinations [@Torok.2017]. This means a close adaptation of the substrate to the niche of the target species or of the seed mixtures to the characteristics of the chosen substrate. Suitable substrates should reduce habitat filtering, while the seed mixture should minimize the competitive exclusion of unsuitable species. Modifying seed mixtures to match specific site conditions could be based on functional plant traits [@Funk.2008; @Laughlin.2014; @Balazs.2020], although this is not easy to implement [@merchant2022; @Bauer.2022]. This challenge is particularly interesting on artificial substrates that are used in urban areas [@Bauer.2022], quarries [@chenot-lescure2022], or on dikes [@Liebrand.1996].

Dikes are promising sites for the restoration of species-rich grasslands because they can increase habitat area and connectivity of semi-natural grasslands and therefore significantly contribute to biodiversity conservation in agricultural landscapes [@Batori.2020]. Steep slopes with different exposition, contrasting substrate layers and dense swards for erosion protection characterize dikes [@Batori.2016; @berendse2015; @husicka2003]. Dikes can reconcile several ecosystem functions like dike security with biodiversity [@teixeira2022], which can be fostered by an adapted seed--substrate combination.

The aim of this study is to identify the best combinations of seed mixtures and substrates for vital and species-rich grasslands on north- and south-exposed dike slopes. An experiment was set up to test agricultural soil with different amounts of sand admixture, substrate depths, seed densities, and seed mixture types. We expected a better development of dry grassland in the south exposition with shallow and sandy substrates, and of mesic meadows in north exposition on less sandy and deep substrates. Furthermore for slopes, higher seed densities are recommended for a succesfull establishment [@KleberLerchbaumer.2017], albeit without experimental evidence.

To measure the success of restoration, we chose three indices. First, we observed the persistence of the sown species, which is a key objective of the restoration measure sowing [@Wilsey.2021]. For this purpose, the presence of the sown species was monitored for three consecutive years. Second, we calculated the Favourable Conservation Status (FCS) which distinguishes habitat-characteristic and non-typical derived diversity [@Helm.2015]. Third, the recovery completeness was determined by measuring the successional distance to the reference grasslands [@Rydgren.2019]. To complement these indices, the success of restoration or the difference from desired conditions was evaluated [cf. @Brudvig.2017b]. This was reached by comparing the species composition of the experimental plots with reference sites. We tested following hypotheses after four years of observation:

1.  Establishment of mesic vs dry grassland mixtures is facilitated by site conditions on northern and southern dike slopes, respectively.

2.  Nutrient reduction (by sand addition) and shallow soil increase establishment of dry grassland compared to mesic mixtures.

3.  Reduced soil resources benefit especially target species in comparison to non-target species(, thus increasing conservation value).

4.  Higher seed density increases generally establishment of sown plants and suppresses non-target species.

\clearpage

\newpage

# Materials & Methods

## Experimental design {.unnumbered}

The combinations of seed mixtures and substrates ('seed--substrate combinations') were tested on a dike at the Danube River in SE Germany (Figure \@ref(fig:map); 314 m a.s.l.; WGS84 (lat/lon), 48.83895, 12.88412). The climate of the region is temperate-suboceanic with a mean annual temperature of 8.4 °C and an annual precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@hari2020; @cdc2022temp; @cdc2022prec], as well as three minor floods which did not reach the plots [Appendix A1, @lfu2021waterlevel; @lfu2021discharge]. The substrates consisted of calcareous sand (0--4 mm) and agricultural soil obtained from a nearby dike construction site near the village of Steinkirchen. A big roller mixed both components and an excavator put the substrates in the prepared plots.

The target vegetation types were lowland mesic hay meadows and semi-dry calcareous grassland [EUNIS codes R22, R1A, @Chytry.2020, 'Arrhenatherion elatioris' and 'Cirsio-Brachypodion pinnati' according to the EuroVegChecklist, codes CM01A, DA01B, @mucina2016]. The species pool of hay meadows and dry grasslands consisted of 55 and 58 species, respectively. The seeds were received from a commercial producer of autochthonous seeds (Co. Krimmer, Pulling). From these species pools, 20 species were selected for each plot in a stratified, randomised manner (Appendix A2). Each mixture contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(tab:seedmix)). The community-weighted means (CWM) of functional traits differed between the seed mixture types (Appendix A3). As expected, the hay-meadow mixtures had a higher specific leaf area (SLA), lower seed mass, and higher canopy height than the dry-grassland mixtures. The south-exposed plots were sown in mid-April 2018 and the north exposition 14 days later. Due to the drought, the south exposition was protected by a geotextile consisting of straw chaff (350 g m^-2^) which was removed after two weeks due to unsatisfactory effects for the seedlings. The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by standard deep cuts with hay removal in July 2019 and 2020. In October 2018, *Bromus hordeaceus* was sown as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 3.0 m, vertically oriented, halfway up the dike slopes, distributed over the north and south exposition, and arranged in six blocks (= replicates). The experiment used a split-plot design combined with a randomized complete block design (Figure \@ref(fig:map)). The split plot was created by the two expositions of the dike, where all 24 treatment combinations were tested, i.e., sand admixtures (0, 25, and 50%), soil depths (15 vs. 30 cm), and the two seed mixture types with the normal seed density and an increased density recommended for slopes [4 vs. 8 g m^-2^, @Kiehl.2010; @KleberLerchbaumer.2017].

Under the substrate, a 5 cm thick drainage layer of gravel (0-16 mm) was installed. Soil samples of the three substrates were tested on both expositions by mixing several sub-samples from different plots. The sand admixture changed the soil texture, increased the C/N ratio, reduced calcium carbonate, but did hardly change the pH which were within the weak alkaline range (Table \@ref(tab:substrate)). The pH corresponded to a well-established dike vegetation on the Rhine River [@husicka2003], phosphate and potassium were rather scarce for agricultural soils, but magnesium showed high concentrations [@lfl2022]. The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and the Londo scale was used [@londo1976]. The establishment rates of species are recorded in Appendix A4. In total, 48 species of the species pool of hay meadows (87%) and 46 (79%) of dry grasslands established by 2021, which are rather good ratios [@hedberg2010]. The species established in 31 ± 22% (mean ± SD) of their sown plots. In total, 274 vascular plant species were detected (Appendix A5). The recovery completeness was described by the successional distance which quantifies the distance of a plot to the average reference site in the ordination [*d~jt,0~*, @Rydgren.2019]. Persistence was derived from the 'species losses' component of the temporal beta-diversity index (TBI; 1 − *B*~sor~) which was calculated by comparing the seed mixtures with the respective species composition of each year using Sørensen dissimilarity [@legendre2019]. The Favourable Conservation Status (FCS) is the ratio of characteristic and derived diversity, in our case species richness. Characteristic diversity consists of species which belong to a habitat-specific species pool and derived diversity consists of all other species. The habitat-specific species pool consisted of all sown species and other typical species of mesic and dry grasslands [Appendix A5, @Helm.2015].

Reference vegetation surveys were extracted from sPlotOpen [@Sabatini.2021] and our own surveys on the Danube dikes in the surroundings [@bauer2023]. We selected six calcareous grassland plots [EUNIS code R1A, @Chytry.2020] within SE Germany from sPlotOpen and 98 plots of our own survey which included also lowland hay meadows (R22) and as a negative reference ruderal, dry and anthropogenic vegetation (V38).

## Statistical analyses {.unnumbered}

To measure the effects of the treatments on our three response variables, we calculated Bayesian linear mixed-effects models (BLMM) with the random effect plot nested in block with the Cauchy prior [see @lemoine2019]. Furthermore, we included as a fixed effect the botanists recording a certain plot. For the simple effects of the treatments, we chose plausible weakly informative priors. To evaluate the influence of the priors, prior predictive checks and models with non-informative priors were calculated. For the computation, we used four chains, a thinning rate of two, 5,000 iterations for warm-up, and 10,000 in total. We used the Markov Chain Monte Carlo method (MCMC) with the no-U-turn Sampler (NUTS), which is based on the Hamiltonian Monte Carlo approach (HMC). For evaluating the computation, the convergence of the four chains was checked using trace plots and evaluating R-hat values, and MCMC chain resolution by the effective sampling size (ESS). Posterior predictive checks were done with Kernel density estimates histograms of statistics skew and leave-one-out (LOO) cross-validation [see @gabry2019]. Finally, the models were compared with the Bayes factor (BF) and Bayesian *R*² values [@gelman2019]. A non-metric multidimensional scaling ordination (NMDS) with Sørensen dissimilarity was used to visualize variation in species composition in space and time (2D-stress 0.21). Seven species were excluded because they had an accumulated cover over all plots of less than 0.5%. Finally, 343 species were included in the ordination.

Data, code and the entire model specifications and evaluations are stored on GitHub and presented in an easily accessible document for scrolling through `r word_spec("(include Zenodo reference)")`.<!--# Cite Zenodo--> There, the sections are referenced to the Bayesian analysis reporting guidelines [BARG, @kruschke2021]. All analyses were performed in R [Version 4.2.2, @base] with the functions 'brm' from the package 'brms' [@brms] for model calculation, several functions from 'brms' and 'bayesplot' [@bayesplot] for model evaluation, and 'metaMDS' from 'vegan' for the ordination [@vegan].

\clearpage

\newpage

# Results

## Hay meadows on north exposition closer to reference

The ordination visualizes the species composition of seed mixtures and the development of the plots over time (Figure \@ref(fig:nmds)). The NMDS confirmed that the seed mixtures were variable, but distinctive for hay meadows and dry grasslands and ascertained the intended direction of the vegetation development. As one exception, hay meadows in south exposition did not develop towards their seed-mixture compositions.

The reference sites had a larger variation than the seed mixtures and were close to the seed mixtures but hardly overlapped (Figure \@ref(fig:nmds)). The positions of the reference sites were shifted to the left in comparison to the seed mixtures, which means in the direction of early successional states. Nonetheless, they still differed from the negative references of ruderal vegetation. Negative references were only available on the south exposition and they were located in the NMDS between the positive reference sites and the state of restored plots in 2021. Hay meadow-seed mixtures led to a closer development to hay meadow references than dry grasslands to their references (Figure \@ref(fig:effects)A, \@ref(fig:interactions)A). This was especially the case in north exposition (Figure \@ref(fig:nmds)).

## Weak effects of restoration treatments

We could identify a statistically clear positive effect of the sand admixture on the persistence of sown species and on the recovery rate (Figure \@ref(fig:effects)). The posterior distributions are also shown in the interaction plots that separate exposition and survey year (Figures \@ref(fig:interactions)). For all three response variables, the vegetation developed strongly positive after one year, but the recovery rate slowed down in the following years. Both expositions had similar trends but for all responses, the values were clearly lower in south exposition, e.g., persistence values were on average more than 46% higher in north exposition (Figure \@ref(fig:interactions)B). The interactions of restoration treatments were neither clear nor strong (Figures \@ref(fig:interactions)). Persistence of both seed mixture types was slightly positively affected by sand admixture on the north exposition (+ 6--7 ± 4%, Figure \@ref(fig:interactions)B). \clearpage \newpage

# Discussion

The vegetation of the dike experiment developed in the desired direction, but with a decelerating recovery rate. A significantly worse effect during establishment phase remained over the complete observation period in south compared to north exposition. All in all, it was hardly possible to identify any best seed--substrate combinations, although the hay meadows developed closer to their reference on the north exposition than the dry grasslands.

## The success of the restoration approaches

The seed mixtures and their positive reference sites were similar but hardly overlapped (Figure \@ref(fig:nmds)). The position on the NMDS suggests that the seed mixture represents a later successional stage than the references. The NMDS shows a slightly better adaptation of hay meadows to the north exposition than of dry grasslands (Figure \@ref(fig:interactions)A). This can be expected from the requirements of hay meadows for mesic conditions which can be provided on north-exposed dike slopes [@Batori.2020; @Oberdorfer.1993]. On the southern exposition, the plots of hay meadows developed rather to dry grassland references which indicates an inefficient restoration by using a non-adapted seed mixture.

The vegetation developed generally in the desired direction but was still distinct from positive references and seed mixtures after four years. In the south exposition, the plots were rather similar to the negative reference of dry, ruderal vegetation. The gap between goal and restoration outcome was also shown for other sowing experiments or restorations [@mitchley2012; @Engst.2016; @Kaulfuss.2022] or for dike vegetation compared with semi-natural references [@Batori.2016]. This result is not surprising since the 'recovery debt' is a general phenomenon of grassland restoration [@MorenoMateos.2017; @jones2018].

## General effects of treatments and exposition

Restoration on agricultural soils can have limited success due to high nutrient loads [@walker2004] but mixing with a mineral component need not necessarily improve the outcome [@chenot-lescure2022]. Similarly to that study in France, sand admixture reduced nutrient loads and led to higher persistence of sown species while a 50% admixture did not further increase the effect. In addition, the effect only appeared in north exposition and the effect size of about 6% in the fourth year after restoration was rather small. The Favourable Conservation Status (FCS) was hardly affected by the sand admixture, which corresponds to the experiment in a quarry [@chenot-lescure2022]. Substrate depth did not significantly affect persistence or FCS, similarly to earlier studies [@baer2004]. Larger differences in soil depths might be necessary to observe negative effects by thicker substrate layers as was shown for prairies [@dornbush2010]. Seed density had also no clear effect on persistence and FCS which fits the results of @Kaulfuss.2022, who found that a certain amount of seeds is necessary for a successful establishment of target species, but higher densities do not further improve the outcome, but rather have a slightly negative effect.

The vegetation in south exposition had a different species composition, which confirms the findings of @Batori.2016 in Hungary. Here, the differences might be due to some unintended methodical reasons: the geotextile, which had been implemented on the southern slope, was removed after two weeks. This led to the drying of many seedlings which was amplified by the intense drought of 2018 and 2019 [@hari2020]. This drought could have decreased germination and the establishment of sown species, also generally speaking, because a moist spring with a mesic temperature enhances establishment [@Larson.2021; @orrock2023]. The interesting point here is the lasting negative effect on persistence and FCS on the southern slope which suggests a legacy effect by adverse weather conditions after sowing as it was observed by other studies [@Stuble.2017; @Groves.2020]. The problems during the establishment phase might have led to an alternative trajectory [@suding2004] and could probably only be met by re-sowing. The drying of seedlings might also have levelled the distinction of the seed mixture types on the south exposition.

## No interaction effect of seed--substrate combinations

Our aim was to identify perfect seed--substrate combinations regarding restoration effectiveness and biodiversity. For evaluating effectiveness, we measured the persistence of the sown species and for fostering plant biodiversity in general, we measured the FCS. However, we could not identify an interaction effect for neither one of the indices. We would have expected a better performance of hay-meadow seed mixtures with lower sand admixture and for dry grasslands with higher sand admixture. Our results suggest that, at least after four years, the substrate conditions are within the range of both seed mixture types (hay meadows, dry grasslands). Although, both types are clearly phytosociologically and functionally distinct, they are still relatively closely related because they contain shared species and could develop under similar site conditions with modified sub-associations [Appendix A3, @husicka2003; @Oberdorfer.1993]. Other grassland studies could identify more or less clear interactions of opposing habitat preferences or functional traits along the gradients of productivity, moisture, and nutrients [@Zirbel.2020; @Freitag.2021; @Kaulfuss.2022]. However, these studies did not work with an experimental set up of different seed--substrate combinations, but analysed the result of habitat and biotic filtering after one, five and about 15 years, respectively. Furthermore, the non-existence of ideal combinations could be explained by priority effects which means that the species of the imperfect-adapted seed mixture type could establish earlier and pre-empted the available niches for the species of related habitat types [@Fukami.2015].

\clearpage

\newpage

# Conclusions

Our results suggest that adapted seed mixtures can increase restoration efficiency by sowing hay meadows in the north but not necessarily in the south exposition of dikes. Furthermore, the reduction of the nutrient load through sand admixture has a clearly positive effect, albeit the effect is small. The question remains if sand admixture is the most economical restoration measure to promote diversity on dikes. Increasing seed density on dike slopes does not appear to be necessary, and soil depths of 30 cm are not adverse compared to 15 cm thick substrates. There were no perfect seed--substrate combinations and thus we conclude that a variation of seed mixture types and different substrates along restoration sections would promote biodiversity better than a single uniform solution [@bauer2023; @Holl.2022]. The strong and lasting negative effects of the drought might require re-sowing. To close the recovery debt, the management adaptation might be promising since this is a crucial factor beside the restoration approach and the site characteristics [@Grman.2013; @tölgyesi2021]. For example, the start of sheep grazing on the experimental plots, which already exists in the surroundings, will modify the disturbance regime and improve dispersal. All in all, our results support the finding that restored dike grasslands can promote biodiversity in agricultural landscapes [@Batori.2020]. However, the recovery debt highlights the fact that restored grasslands cannot substitute old-growth grasslands [@nerlekar2020; @Buisson.2022].

\clearpage

\newpage

# Acknowledgements {.unnumbered}

We would like to thank our project partners Dr. Markus Fischer, Frank Schuster, and Christoph Schwahn (WIGES GmbH) as well as Klaus Rachl and Stefan Radlmair (Government of Lower Bavaria) for numerous discussions on restoration and management of dike grasslands. Field work was supported by Clemens Berger and Uwe Kleber-Lerchbaumer (Wasserwirtschaftsamt Deggendorf). We thank Holger Paetsch, Simon Reith, Anna Ritter, Jakob Strak, Leonardo H. Teixeira, and Linda Weggler for assisting with the field surveys or soil analyses in 2018--2020. The German Federal Environmental Foundation (DBU) supported MB with a doctoral scholarship.

# Author contribution {.unnumbered}

JH and JK designed the experiment. JH did the surveys in the years 2018--2020 and MB in 2019 and 2021. MB did the analyses and wrote the manuscript. JK and JH critically revised the manuscript.

# Open research {#sec-open-research .unnumbered}

Data and code are stored on Zenodo (Bauer et al. 2023) <!--# Update citation after upload -->Model evaluation is stored on GitHub: <https://github.com/markus1bauer/2023_danube_dike_experiment/tree/main/markdown>

# Funding {.unnumbered}

MB was funded by a doctoral scholarship of the German Federal Environmental Foundation (DBU) (No. 20021/698). The establishment of the experiment and the vegetation surveys were financed by the WIGES GmbH in the years 2018--2020 (No. 80 002 312).

\clearpage

\newpage

# References {.unnumbered}

::: {#refs}
:::

\clearpage

\newpage

<!---BLOCK_LANDSCAPE_START--->

# Tables {.unnumbered}

## Table 1 {.unnumbered}

```{r tab.id = "seedmix", ft.align = "left", tab.cap = "Each plot received an individual set of twenty species with some restrictions to the number of species per functional group. The total species pool for hay meadows was 55 and for dry grassland 58. In total, plots received 16 or 32 g of seeds. All individual seed mixtures are stored in Appendix A2."}
read_csv(here::here("data", "raw", "data_raw_table_1_seedmix.csv"),
         show_col_types = FALSE) %>%
  slice(-1) %>%
  flextable() %>%
  merge_at(part = "header", j = 2:3) %>%
  align(align = "center", part = "all", j = -1) %>%
  add_header_row(
    top = FALSE,
    values = c("", "Hay meadow", "Dry grassland", "", "", "")
    ) %>%
  add_header_row(top = FALSE, values = c("", "#", "#", "#", "wt%", "wt%")) %>%
  bold(part = "header", i = 1) %>%
  border_inner_h(part = "header", border = fp_border(width = 1)) %>%
  autofit()
```

\clearpage

\newpage

## Table 2 {.unnumbered}

Characteristics of the substrates used for the sowing experiment on river dikes. Soil samples of the three substrates were analyzed for the fraction \< 2 mm. The soil texture was classified according to the 'Bodenkundliche Kartieranleitung' [@BodenkundlKartieranl.2005]. The pH was measured in CaCl~2~ solution. Plant available phosphorus and potassium were measured in a calcium acetate-lactate extract and magnesium in a CaCl~2~ extract. For calculating CaCO~3~, a sub-sample was annealed at 550 °C and the measured C amount multiplied with 8.33. To calculate total N and the C/N ratio, a sub-sample was incinerated at 1000 °C. Lt3 = medium clayey loam; Ls4 = strong sandy loam; Sl3 = medium loamy sand; Sl4 = strong loamy sand.

```{r tab.id = "substrate", ft.align = "left", tab.cap = ""}
read_csv(here::here("data", "raw", "data_raw_table_3_substrate.csv"),
         show_col_types = FALSE) %>%
  select(-starts_with("lfl")) %>%
  slice(-1) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(
    top = FALSE,
    values = read_csv(
      here::here("data", "raw", "data_raw_table_3_substrate.csv"),
                 show_col_types = FALSE) %>%
      select(-starts_with("lfl")) %>%
      slice(1)
    ) %>%
  border_inner_h(part = "header", border = fp_border(width = 1)) %>%
  align(align = "center", part = "all", j = -1) %>%
  compose(
    i = 1, j = 10, part = "header",
    value = as_paragraph("P", as_sub("2"), "O", as_sub("5"))
    ) %>%
  compose(
    i = 1, j = 11, part = "header",
    value = as_paragraph("K", as_sub("2"), "O")
    ) %>%
  compose(
    i = 1, j = 12, part = "header",
    value = as_paragraph("Mg", as_sup("2+"))
    ) %>%
  compose(
    i = 1, j = 14, part = "header",
    value = as_paragraph("CaCO", as_sub("3"))
    ) %>%
  compose(
    i = 2, j = 10, part = "header",
    value = as_paragraph("mg 100 g", as_sup("−1"))
    ) %>%
  compose(
    i = 2, j = 11, part = "header",
    value = as_paragraph("mg 100 g", as_sup("−1"))
    ) %>%
  compose(
    i = 2, j = 12, part = "header",
    value = as_paragraph("mg 100 g", as_sup("−1"))
    ) %>%
  align(align = "left", part = "footer") %>%
  set_table_properties(layout = "autofit", width = 1)
```

\newpage

<!---BLOCK_LANDSCAPE_STOP--->

# Figures {.unnumbered}

## Figure 1 {.unnumbered}

```{r fig.id = "map", fig.asp = 1.2854, fig.width = 5, fig.cap = ""}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_1_map.png")
  )
```

Local setting and design of the multifactorial experiment on grassland sowing at dikes. The experiment was located on a dike at the Danube River in SE Germany. The 288 plots were allocated in six blocks (white squares on the upper photo) and on the north and south slope (central photo) [Both aerial photos: @bayerischevermessungsverwaltung2023]. Four treatments were conducted: sand admixture, substrate depth, seed density, and seed mixture types H and D (hay meadows, dry grasslands). The western half of a block had a shallow substrate depth and within this, half of the substrates had different sand admixtures. The photo on the bottom shows the northern slope of one block in 2021, four years after sowing (Photo: Markus Bauer).

\clearpage

\newpage

## Figure 2

```{r fig.id = "nmds", fig.asp = 0.9696, fig.width = 6.49, fig.cap = ""}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_2_nmds_800dpi_16.5x16cm.tiff")
  )
```

The species composition of sown experimental plots on a river dike over time and in comparison with reference sites and the seed mixtures. Both expositions and both seed mixture types are shown in separated panels. The NMDS was based on the Sørensen dissimilarity and data of 288 plots observed over four years after sowing in 2018 (circles). These experimental plots were compared with the seed mixtures (black squares) and 98 positive and negative reference plots (filled symbols) from older dike grasslands in the surroundings [@bauer2023] and six plots from sPlotOpen [@Sabatini.2021]. The ellipses show the standard error of the groups. 2D-stress: 0.21.

\clearpage

\newpage

## Figure 3 {.unnumbered}

```{r fig.id = "effects", fig.asp = 0.3030, fig.width = 6.49, fig.cap = ""}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_3_800dpi_16.5x5cm.tiff")
  )
```

Effects of treatments on the development of sown grassland communities at a river dike. The posterior density distributions (grey) is calculated over all four surveyed years and both expositions. Shown are the medians, 66% and 95% credible intervals, which were derived from a Bayesian linear mixed-effects model (BLMM). (A) The persistence of sown species, (B) the Favourable Conservation Status (FCS) and (C) the recovery completeness compared to reference sites. The FCS is the ratio of target species to non-target species. Note that the zero lines indicate that both levels have equal values. This means, e.g., that hay meadows are closer to their reference than dry grasslands (C).

\clearpage

\newpage

## Figure 4 {.unnumbered}

```{r fig.id = "interactions", fig.asp = 1.2121, fig.width = 6.49, fig.cap = ""}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_4_800dpi_16.5x20cm.tiff")
  )
```

The development sown grassland communities at a river dike. The plots had substrates with different sand admixtures and were sown with two different seed mixture types. Three indices are evaluated. (A) Recovery completeness (*d~jt~*~,0~): the zero lines indicate the mean position of the reference sites for each habitat type on the NMDS axis 1 (Figure \@ref(fig:nmds)). The grey area marks the standard deviation of the position of the reference sites. (B) Persistence of sown species: losses component of the temporal beta-diversity index (1 − *B*~sor~) (C) Favourable Conservation Status (FCS): the zero line indicates that target and non-target species are balanced. Positive values indicate that there are more target species. Shown are the medians and 95% credible intervals of the posterior distributions, which were derived from a Bayesian linear mixed-effects model (BLMM).

\clearpage

\newpage

# Session Info

```{r sessioninfo}
sessionInfo()
```